---
title: "Ex. 2 Analysis"
output: html_document
date: "2023-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('helper functions.R')
```

```{r}
#reading in the data
data.many_csv <- read.csv("Study 3 (Study 2 Rep) Final Data - Many Condition.csv")
data.some_csv <- read.csv("Study 3 (Study 2 Rep) Final Data - Some Condition.csv")
```


```{r}

#removing participants who failed the catch trials
data.many_csv <- data.many_csv %>% filter(catch.1 == "x" & (catch.2 =="x" | catch.2 == "skipped") & (catch.3 =="x" | catch.3 == "skipped") & (catch.4 =="x" | catch.4 == "skipped") & (catch.5 =="x" | catch.5 == "skipped") & (catch.6 =="x" | catch.6 == "skipped") & (catch.7 =="x" | catch.7 == "skipped") & (catch.8 =="x" | catch.8 == "skipped"))

data.many <- data.many_csv %>% pivot_longer(c(-Subject.ID, -Age, -Gender))

data.many$condition = "many"

#adding a column for number of candies and phase
data.post.many <- data.many %>% filter(grepl("^X",name))
data.post.many <- data.post.many %>% mutate(nCandies = sub("X([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.post.many$phase <- "post"

data.pre.many <- data.many %>% filter(grepl("^PE.[0-9]",name))
data.pre.many <- data.pre.many %>% mutate(nCandies = sub("PE.([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.pre.many$phase <- "pre"

#adding numeric values: 1 if response is many, 0 if response is some; filter out any other response
data.many.full <- rbind(data.pre.many,data.post.many)
data.many.full <- data.many.full %>% mutate(num_response = case_when(value=="many"~1, value == "some"~0))
data.many.full <- data.many.full %>% filter(!is.na(num_response))

#finding means and confidence intervals
many_condition_means <- data.many.full %>%
  group_by(Subject.ID, phase, nCandies) %>% 
  summarise(subject_mean = mean(num_response)) %>%
  group_by(phase, nCandies) %>%
  summarise(mean= mean(subject_mean),
         ci_high = mean+ ci.high(subject_mean), 
         ci_low = mean- ci.low(subject_mean)
         )
  
many_condition_means$condition <- "Many-biased Condition"

many_condition_plot <- ggplot(many_condition_means, aes(x = nCandies, y = mean, color=phase)) + geom_point() + geom_line() + geom_errorbar(ymin = many_condition_means$ci_low, ymax= many_condition_means$ci_high) + labs(x = "Number of Candies out of 10", y= "Proportion Choosing Many", title = "Many Condition (e.g., heard the speaker say many for 5/10)") + ylim(0,1)


many_condition_plot
```


```{r}
#removing participants who failed the catch trials
data.some_csv <- data.some_csv %>% filter(catch.1 == "x" & (catch.2 =="x" | catch.2 == "skipped") & (catch.3 =="x" | catch.3 == "skipped") & (catch.4 =="x" | catch.4 == "skipped") & (catch.5 =="x" | catch.5 == "skipped") & (catch.6 =="x" | catch.6 == "skipped") & (catch.7 =="x" | catch.7 == "skipped") & (catch.8 =="x" | catch.8 == "skipped"))

data.some <- data.some_csv %>% pivot_longer(c(-Subject.ID, -Age, -Gender))
data.some $condition = "some"

#adding a column for number of candies and phase
data.post.some <- data.some %>% filter(grepl("^X",name))
data.post.some <- data.post.some %>% mutate(nCandies = sub("X([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.post.some$phase <- "post"

data.pre.some <- data.some %>% filter(grepl("^PE.[0-9]",name))
data.pre.some <- data.pre.some %>% mutate(nCandies = sub("PE.([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.pre.some$phase <- "pre"


data.some.full <- rbind(data.pre.some,data.post.some)

#adding numeric values: 1 if response is many, 0 if response is some; filter out any other response
data.some.full <- data.some.full %>% mutate(num_response = case_when(value == "many"~1, value == "some" ~0))
data.some.full <- data.some.full %>% filter(!is.na(num_response))

#combining many condition data and some condition data
data_total <- rbind(data.some.full, data.many.full)

#centering number of candies
data_total$n_centered <- data_total$nCandies-5


#finding means and confidence intervals
some_condition_means <- data.some.full %>%
  group_by(Subject.ID, phase, nCandies) %>% 
  summarise(subj_mean = mean(num_response)) %>%
  group_by(phase, nCandies) %>%
  summarise(mean = mean(subj_mean),
         ci_high = mean + ci.high(subj_mean), 
         ci_low = mean - ci.low(subj_mean)
         )

some_condition_plot <- ggplot(some_condition_means, aes(x = nCandies, y = mean, color=phase)) + 
    geom_point() + geom_line() + geom_errorbar(ymin = some_condition_means$ci_low, ymax= some_condition_means$ci_high) + 
  labs(x = "Number of Candies out of 10", y= "Proportion Choosing Many", title = "Some Condition (e.g., heard the speaker say some for proportions of 5/10)") + ylim(0,1)

some_condition_plot

#creating a data frame to plot both the change in some responses and many responses (change in some and many are basically mirror image after removing "other" responses, so it is not super informative, but it makes a nice looking plot)
data.some.full_some <- data.some.full %>% mutate(num_response_reversed = case_when(num_response == 0 ~ 1,
                                                                                   .default = 0))
some_condition_means_some = data.some.full_some %>%
  group_by(Subject.ID, phase, nCandies) %>% 
  summarise(participant_mean = mean(num_response_reversed)) %>%
  group_by(phase, nCandies) %>%
  summarise(mean= mean(participant_mean),
         ci_high = mean+ ci.high(participant_mean), 
         ci_low = mean- ci.low(participant_mean)
         )
  
some_condition_means_some$response <- "some"
some_condition_means$response <- "many"
some_condition_means_some$condition <- "Some-biased Condition"
some_condition_means$condition <- "Some-biased Condition"
some_total <- rbind(some_condition_means_some, some_condition_means)

data.many.full_some <- data.many.full %>% mutate(num_response_reversed = case_when(num_response == 0 ~ 1,
                                                                                   .default = 0))
many_condition_means_some = data.many.full_some %>%
  group_by(Subject.ID, phase, nCandies) %>% 
  summarise(participant_mean = mean(num_response_reversed)) %>%
  group_by(phase, nCandies) %>%
  summarise(mean = mean(participant_mean),
         ci_high = mean + ci.high(participant_mean), 
         ci_low = mean - ci.low(participant_mean)
         )

many_condition_means_some$response <- "some"
many_condition_means$response <- "many"
many_total <- rbind(many_condition_means_some, many_condition_means)
many_total$condition <- "Many-biased Condition"

total_means_many <- rbind(many_condition_means, some_condition_means)
total_means <- rbind(some_total, many_total)

change_in_many_response_plot <- ggplot(total_means_many, aes(x = nCandies, y = mean, color=condition)) + 
    geom_point() + geom_line() + geom_errorbar(ymin = total_means_many$ci_low, ymax= total_means_many$ci_high) +   
  facet_wrap(~phase) + labs(x = "Number of Candies out of 10", y = "Proportion Choosing Many", title = "Mean Choice of Many over Some by condition and phase") + ylim(0,1)

change_in_many_response_plot

total_change_plot <- ggplot(total_means, aes(x = nCandies, y = mean, color=response, lty=phase)) + 
    geom_point() + geom_line() + geom_errorbar(ymin = total_means$ci_low, ymax= total_means$ci_high) + facet_wrap(~condition) +
    labs(x = "Number of Candies out of 10", y = "Proportion of each choice", title = "Mean Change in Choice of Some and Many by phase and condition") + ylim(0,1)

total_change_plot
```


```{r}
#pre-registered model
prepost_model <- glmer(num_response ~ n_centered * condition * phase  + (1 | Subject.ID), 
      data = data_total, 
      family = "binomial")

summary(prepost_model)
```


There doesn't seem to be much of an effect in the "some" condition, but maybe this is because for 5/10 candies, participants were already expecting the speaker to say "some" (>80%), so the exposure did not provide any new information. Below, I repeat the above analysis and recreate the plots, but I remove any participant who exclusively answered "some" for 5/10 (aka anyone who was already at ceiling so could not have adapted for the target proportion).
```{r}
data_many_exlude_ceiling <- data.many_csv %>% filter(PE.5 == "some" | PE.5.1 == "some" | PE.5.2 == "some")

data_many_exlude_ceiling <- data_many_exlude_ceiling %>% pivot_longer(c(-Subject.ID, -Age, -Gender))


data_many_exlude_ceiling$condition = "many"

#adding a column for number of candies and phase
data.post.many <- data_many_exlude_ceiling %>% filter(grepl("^X",name))
data.post.many <- data.post.many %>% mutate(nCandies = sub("X([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.post.many$phase <- "post"

data.pre.many <- data_many_exlude_ceiling %>% filter(grepl("^PE.[0-9]",name))
data.pre.many <- data.pre.many %>% mutate(nCandies = sub("PE.([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.pre.many$phase <- "pre"

#adding numeric values: 1 if response is many, 0 if response is some; filter out any other response
data.many.full <- rbind(data.pre.many,data.post.many)
data.many.full <- data.many.full %>% mutate(num_response = case_when(value=="many"~1, value == "some"~0))
data.many.full <- data.many.full %>% filter(!is.na(num_response))

#finding means and confidence intervals
many_condition_means <- data.many.full %>%
  group_by(Subject.ID, phase, nCandies) %>% 
  summarise(subject_mean = mean(num_response)) %>%
  group_by(phase, nCandies) %>%
  summarise(mean= mean(subject_mean),
         ci_high = mean+ ci.high(subject_mean), 
         ci_low = mean- ci.low(subject_mean)
         )
  
many_condition_means$condition <- "many"

many_condition_exclude_ceiling_plot <- ggplot(many_condition_means, aes(x = nCandies, y = mean, color=phase)) + geom_point() + geom_line() + geom_errorbar(ymin = many_condition_means$ci_low, ymax= many_condition_means$ci_high) + labs(x = "Number of Candies out of 10", y= "Proportion Choosing Many", title = "Many Condition (excludes participants who only answered many for 5/10)")

many_condition_exclude_ceiling_plot
```


```{r}
data_some_exlude_ceiling <- data.some_csv %>% filter(PE.5 == "many" | PE.5.1 == "many" | PE.5.2 == "many")

data_some_exlude_ceiling <- data_some_exlude_ceiling %>% pivot_longer(c(-Subject.ID, -Age, -Gender))

data_some_exlude_ceiling$condition = "some"

#adding a column for number of candies and phase
data.post.some <- data_some_exlude_ceiling %>% filter(grepl("^X",name))
data.post.some <- data.post.some %>% mutate(nCandies = sub("X([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.post.some$phase <- "post"

data.pre.some <- data_some_exlude_ceiling %>% filter(grepl("^PE.[0-9]",name))
data.pre.some <- data.pre.some %>% mutate(nCandies = sub("PE.([0-9]).*", "\\1", name, perl = TRUE)) %>% mutate(nCandies=as.numeric(nCandies))
data.pre.some$phase <- "pre"

#adding numeric values: 1 if response is many, 0 if response is some; filter out any other response
data.some.full <- rbind(data.pre.some,data.post.some)
data.some.full <- data.some.full %>% mutate(num_response = case_when(value=="many"~1, value == "some"~0))
data.some.full <- data.some.full %>% filter(!is.na(num_response))

#finding means and confidence intervals
some_condition_means <- data.some.full %>%
  group_by(Subject.ID, phase, nCandies) %>% 
  summarise(subject_mean = mean(num_response)) %>%
  group_by(phase, nCandies) %>%
  summarise(mean= mean(subject_mean),
         ci_high = mean+ ci.high(subject_mean), 
         ci_low = mean- ci.low(subject_mean)
         )
  
some_condition_means$condition <- "some"

some_condition_exclude_ceiling_plot <- ggplot(some_condition_means, aes(x = nCandies, y = mean, color=phase)) + geom_point() + geom_line() + geom_errorbar(ymin = some_condition_means$ci_low, ymax= some_condition_means$ci_high) + labs(x = "Number of Candies out of 10", y= "Proportion Choosing Many", title = "Some Condition (excludes participants who only answered some for 5/10)")


some_condition_exclude_ceiling_plot

total_exclusion_data <- rbind(data.some.full, data.many.full)
```

Re-running the model with the trimmed data
```{r}
total_exclusion_data$n_centered <- total_exclusion_data$nCandies-5

excluded_model <- glmer(num_response ~ n_centered * condition * phase  + (1 | Subject.ID), 
      data = total_exclusion_data, 
      family = "binomial")

summary(excluded_model)
```


